---
title: "cs1Xiaojun"
output: pdf_document
---

```{r}
library(ggplot2)
library(dplyr)
library(MASS)
library(ggpubr)
library(monomvn)
Longnecker <- as.data.frame(readRDS("Longnecker.rds"))
head(Longnecker)
Longnecker$center<- as.factor(as.character(Longnecker$center))
Longnecker$center <- relevel(Longnecker$center, ref = "5")
Longnecker[Longnecker$gestational_age > 46, "gestational_age"] <- 46 # Truncate at age == 46
data = Longnecker
```

## about 93% missing for albumin, 20% missing for scores
## if we consider gestational_age > 42 as measurement error, we have about 200 observations
- not a good idea to delete them
- might think about setting those values as 42
## Correlation between variables
- Large association between PCBs expect the first two concentration values
- Not able to find great association between gestational age and other variables
## Should use log pcb to make it normal
## pcb 074, 170 and gestational age. It seems the association becomes more distinguishable among race as the concentration increases
## center 15 and 37 tend to have high premature ratio, and we find that those centers have mainly black people
## unbalanced number between white and black if the maternal age is <20
## center 5 and 66 have more older mother

```{r}
### missing data for each variable
colSums(is.na(Longnecker))
colSums(is.na(Longnecker))/nrow(Longnecker)

### Gestational_age
hist(Longnecker$gestational_age)
### above 42 might be measurement error
sum(Longnecker$gestational_age > 46)

### Correlation between variables
data_corr<- Longnecker %>%
  dplyr::select(-c(albumin, score_education, score_income, score_occupation))
data_corr$race<- as.numeric(data_corr$race)
data_corr_age<- data_corr[data_corr$gestational_age <=42, ]

M=cor(as.matrix(na.omit(data_corr)))
corrplot.mixed(M,number.cex=.6)

M=cor(as.matrix(na.omit(data_corr_age))) ### correlation between gestational age and others becomes higher when we deleted the measurement error
corrplot.mixed(M,number.cex=.6)

### PCB
hist(Longnecker$pcb_028)
hist(log(Longnecker$pcb_028))
hist(Longnecker$pcb_052)
hist(log(Longnecker$pcb_052))
hist(Longnecker$pcb_074)
hist(log(Longnecker$pcb_074))
hist(Longnecker$pcb_170)
hist(log(Longnecker$pcb_170))

### race
table(Longnecker$race)

#data_age<- data[data$gestational_age <=46, ]
### Three more variables
qplot(x = cholesterol, y = gestational_age, data=data_age, color = factor(smoking_status))
qplot(x = cholesterol, y = gestational_age, data=data, color = factor(smoking_status))

qplot(x = triglycerides, y = gestational_age, data=data_age, color = factor(smoking_status))
qplot(x = dde, y = gestational_age, data=data_age, color = factor(smoking_status))

qplot(x = dde, y = gestational_age, data=data_age, color = race)
qplot(x = pcb_028, y = gestational_age, data=data_age, color = race)
qplot(x = pcb_052, y = gestational_age, data=data_age, color = race)
### seems to have some relationship 
qplot(x = pcb_074, y = gestational_age, data=data_age, color = race)
qplot(x = pcb_170, y = gestational_age, data=data_age, color = race)
qplot(x = pcb_194, y = gestational_age, data=data_age, color = race)

### center 15 and 37
table(data_age$center)
boxplot(gestational_age ~ center, data=data_age, col="orange", border="brown")
premature<- c()
premature_ratio<- c()
for (i in unique(data_age$center)){
   pre <- sum(data_age[data_age$center == i, ]$gestational_age <37)
   premature<- c(premature, pre)
   pre_ratio<- pre/sum(data_age$center == i)
   premature_ratio<- c(premature_ratio, pre_ratio)
}

data.frame(ratio= premature_ratio, center= unique(data_age$center))

c15<- data_age %>% filter(center == 15)
qplot(x = dde, y = gestational_age, data = c15, color = race, geom= c("point", "line"))

c37<- data_age %>% filter(center == 37)
qplot(x = dde, y = gestational_age, data = c37, color = race, geom= c("point", "line"))

c10<- data_age %>% filter(center == 10)
qplot(x = dde, y = gestational_age, data = c10, color = race, geom= c("point", "line"))

ggplot(data_age, aes(gestational_age, fill = race)) + geom_density(alpha = 0.5)

table(data_age$center, data_age$race)
table(data_age$maternal_age, data_age$race) ## unbalanced number between white and black if the maternal age is <20
table(data_age$maternal_age, data_age$center) ### center 5 and 66 have more older mother

boxplot(gestational_age ~ smoking_status, data=data_age, col="orange", border="brown") ### might have a little difference

cor(Longnecker$score_education, Longnecker$gestational_age)
```

```{r}
full.lm<- lm(gestational_age ~ dde + pcb_028 + pcb_052 + pcb_074 + pcb_105 + pcb_118 + pcb_153+ pcb_170 + pcb_138 + pcb_180 + pcb_194 + pcb_203 + triglycerides + race + maternal_age + smoking_status + cholesterol + center, data = data_age)

summary(full.lm)
plot(full.lm, ask=F)
outlierTest(full.lm)

stepAIC(full.lm, direction="both")

full.lm2<- lm(formula = gestational_age ~ dde + pcb_074 + pcb_118 + pcb_170 + pcb_138 + pcb_194 + triglycerides + race + cholesterol, data = data_age)
summary(full.lm2)
```

```{r}
data_age_na<- data_age[-which(is.na(data_age$pcb_028)), ]

library(BAS)
age.bas = bas.lm(gestational_age~ dde + pcb_028 + pcb_052 + pcb_074 + pcb_105 + pcb_118 + pcb_153+ pcb_170 + pcb_138 + pcb_180 + pcb_194 + pcb_203 + race + triglycerides + cholesterol+ maternal_age + smoking_status + center + race*dde  + race*triglycerides + race*cholesterol+ race*maternal_age + race*smoking_status + center*race, data=data_age_na, prior = 'JZS', modelprior=uniform(), method="MCMC", pivot=TRUE,  # for non-full rank case
                   n.models=50000,
                   force.heredity = TRUE)
summary(age.bas)

### pcb 028, 052, 074, 153, 138, 180, 194, 203 could be used in the full model bc we dont want highly correlation covariates in the model

age.bas = bas.lm(gestational_age~ dde + pcb_028 + pcb_052 + pcb_074 + pcb_153+ pcb_138 + pcb_180 + pcb_194 + pcb_203 + race + triglycerides + cholesterol+ maternal_age + smoking_status + center + race*dde  + race*triglycerides + race*cholesterol+ race*maternal_age + race*smoking_status + center*race, data=data_age_na, prior = 'JZS', modelprior=uniform(), method="MCMC", pivot=TRUE,  # for non-full rank case
                   n.models=50000,
                   force.heredity = TRUE)
summary(age.bas)


age.bas = bas.lm(gestational_age~ dde + pcb_028 + pcb_074 + pcb_105 + pcb_118 + pcb_153 + pcb_138 + pcb_180 + race + triglycerides + cholesterol+ maternal_age + smoking_status + center + race*dde  + race*triglycerides + race*cholesterol+ race*maternal_age + race*smoking_status + center*race, data=data_age_na, prior = 'JZS', modelprior=uniform(), method="MCMC", pivot=TRUE,  # for non-full rank case
                   n.models=50000,
                   force.heredity = TRUE)
summary(age.bas)

BMA_coef <- coef(age.bas)
n.models <- 2
BMA_coef <- coef(age.bas, n.models = n.models)
plot(BMA_coef, ask = F)


### dde, race, triglycerides

full.lm<- lm(gestational_age ~ dde + pcb_028 + pcb_052 + pcb_074 + pcb_105 + pcb_118 + pcb_153+ pcb_170 + pcb_138 + pcb_180 + pcb_194 + pcb_203 + race + triglycerides + cholesterol+ maternal_age + smoking_status + center + race*dde + race*triglycerides + race*cholesterol+ race*maternal_age + race*smoking_status + race*center, data = data_age_na)

### remove those have lots of 0s

full.lm<- lm(gestational_age~ dde + pcb_028 + pcb_074 + pcb_105 + pcb_118 + pcb_153 + pcb_138 + pcb_180 + race + triglycerides + cholesterol+ maternal_age + smoking_status + center + race*dde  + race*triglycerides + race*cholesterol+ race*maternal_age + race*smoking_status + center*race, data = data_age_na)

#summary(full.lm)
#plot(full.lm, ask=F)
#outlierTest(full.lm)

#stepAIC(full.lm, direction="both")
X = model.matrix(full.lm)
Y = data_age_na$gestational_age
nrow(X); length(Y);

bhs.fit = bhs(X[,-1], Y, T=1000, normalize=T)

beta.sim = bhs.fit$beta
colnames(beta.sim) = colnames(X)[-1]
quant5 = function(x) {round(quantile(x, c(0.025, 0.5, 0.975)),2)} ## 95% CI
quant10 = function(x) {round(quantile(x, c(0.05, 0.5, 0.95)),2)} ## 90% CI

apply(beta.sim, 2, quant5)
apply(beta.sim, 2, quant10)
### Not sure why mean is always 0
```


```{r}
#library(rjags)
library(R2jags)
full.lm<- lm(gestational_age ~ dde + pcb_028 + pcb_052 + pcb_074 + pcb_105 + pcb_118 + pcb_153+ pcb_170 + pcb_138 + pcb_180 + pcb_194 + pcb_203 + race + triglycerides + cholesterol+ maternal_age + smoking_status + center + race*dde + race*triglycerides + race*cholesterol+ race*maternal_age + race*smoking_status + center*race, data = data_age_na)

Y = data_age_na$gestational_age
X0 = model.matrix(full.lm)
# remove X where the fitted value is NA
coef(full.lm)[is.na(coef(full.lm))] #EE0.01:labBayer  EE0.03:labBayer EE0.01:labZeneca 
namesX0 = colnames(X0)
X0 = X0[,-c(which(namesX0=="raceblack:center15"),
           which(namesX0=="raceother:center15"),
           which(namesX0=="raceother:center45"),
           which(namesX0=="raceblack:center50"),
           which(namesX0=="raceother:center82"))]
X1 = dplyr::select(data.frame(X0),starts_with("center"))
X2 = dplyr::select(data.frame(X0),-starts_with("center"))[,-1]
X = as.matrix(cbind(X1,X2))
model = function(){
  for (i in 1:n){
    Y[i] ~ dnorm(alpha + Xs[i,] %*% beta, phi)
  }
  
  alpha ~ dnorm(0, .000001*phi)
  phi ~ dgamma(.001, .001)
  sigma_L ~ dt(0,1,1)
  phi_L <- 1/((sigma_L^2)+.000001)
  tau ~ dgamma(.5, .5*n)
  sigma <- sqrt(1/phi)
  
  # beta for lab
  for (j in 1:p_center){
    lambda[j] ~ dgamma(a/2, a/2)
    beta[j] ~ dnorm(0, phi*phi_L*lambda[j]) # beta for lab starts from second column
  }
  
  # beta for others
  beta[(p_center+1):p] ~ dmnorm(rep(0,p-p_center), phi*tau*SSX2)
}
set.seed(1)
Xs = scale(X)
data = list(Y=Y, Xs=Xs, SSX2 = t(Xs[,(ncol(X1)+1):ncol(X)])%*%Xs[,(ncol(X1)+1):ncol(X)],
            n=length(Y), p_center=ncol(X1), p=ncol(X), a=9)
output = jags(data, inits=NULL,
              parameters.to.save=c("alpha","beta","sigma","sigma_L","lambda"),
              model=model, n.iter=10000)
sim.matrix = output$BUGSoutput$sims.matrix


n = length(Y)
S = diag(c(apply(X,2,function(x) {var(x)})),nrow=ncol(X))
# Transform back original betas
beta.sim = sim.matrix[,2:57] %*% (solve(S)^0.5) #beta_s = S^0.5*beta
#max(abs(beta.sim %*% (S^0.5) - sim.matrix[,2:57])) # Checkings
# Transform back original alpha
alphaS.sim = sim.matrix[,1]
Xbar = apply(X,2,mean)
alpha.sim = alphaS.sim - beta.sim %*% Xbar
colnames(alpha.sim) = "(Intercept)"
colnames(beta.sim) = colnames(X)

quant5 = function(x) {round(quantile(x, c(0.025, 0.5, 0.975)),2)} ## 95% CI
quant10 = function(x) {round(quantile(x, c(0.05, 0.5, 0.95)),2)} ## 90% CI
apply(beta.sim, 2, quant5)
apply(beta.sim, 2, quant10)

gelman.diag(as.mcmc(output))

### However, mixing is not good for sigma_L, which is the variation between centers. 
```

```{r}
center = dplyr::select(data.frame(beta.sim),starts_with("center"))
center_fit<- cbind(lwr= apply(center, 2, quant5)[1,], fit= apply(center, 2, quant5)[2,], upr=apply(center, 2, quant5)[3,])

ggplot(as.data.frame(center_fit), aes(rownames(center_fit), 
  fit, size=10, group=1, ylim=max(upr)+0.8)) +
  theme_bw(base_size=10)+ 
  geom_point(size=2)+
  geom_line(size=1)+
  geom_errorbar(aes(ymin = lwr, ymax = upr), width = 0.2, size=1)+
  xlab("Lab")+
  ylab("Fitted value")+
  ggtitle("Gestational Age acorss centers with ref=5 (JAGS)")+
  theme(axis.text.x=element_blank())+
  geom_text(aes(label = round(lwr,2), y = lwr), vjust = 1.5, size=3) +
  geom_text(aes(label = round(upr,2), y = upr), vjust = -0.5, size=3) +
  geom_text(aes(label = rownames(center_fit), y = upr+0.5), vjust = -0.5, size=3, col="blue")


pcb = dplyr::select(data.frame(beta.sim),starts_with("pcb_"))
pcb_fit<- cbind(lwr= apply(pcb, 2, quant5)[1,], fit= apply(pcb, 2, quant5)[2,], upr=apply(pcb, 2, quant5)[3,])

ggplot(as.data.frame(pcb_fit), aes(rownames(pcb_fit), 
  fit, size=10, group=1, ylim=max(upr)+0.8)) +
  theme_bw(base_size=10)+ 
  geom_point(size=2)+
  geom_line(size=1)+
  geom_errorbar(aes(ymin = lwr, ymax = upr), width = 0.2, size=1)+
  xlab("Lab")+
  ylab("Fitted value")+
  ggtitle("Gestational Age acorss (JAGS)")+
  theme(axis.text.x=element_blank())+
  geom_text(aes(label = round(lwr,2), y = lwr), vjust = 1.5, size=3) +
  geom_text(aes(label = round(upr,2), y = upr), vjust = -0.5, size=3) +
  geom_text(aes(label = rownames(pcb_fit), y = upr+0.5), vjust = -0.5, size=3, col="blue")

black_interact = dplyr::select(data.frame(beta.sim),starts_with("raceblack"))
black_fit<- cbind(lwr= apply(black_interact, 2, quant5)[1,], fit= apply(black_interact, 2, quant5)[2,], upr=apply(black_interact, 2, quant5)[3,])

ggplot(as.data.frame(black_fit), aes(rownames(black_fit), 
  fit, size=10, group=1, ylim=max(upr)+0.8)) +
  theme_bw(base_size=10)+ 
  geom_point(size=2)+
  geom_line(size=1)+
  geom_errorbar(aes(ymin = lwr, ymax = upr), width = 0.2, size=1)+
  xlab("Lab")+
  ylab("Fitted value")+
  ggtitle("Gestational Age vs. black interaction terms (JAGS)")+
  theme(axis.text.x=element_blank())+
  geom_text(aes(label = round(lwr,2), y = lwr), vjust = 1.5, size=3) +
  geom_text(aes(label = round(upr,2), y = upr), vjust = -0.5, size=3) +
  geom_text(aes(label = rownames(black_fit), y = upr+0.5), vjust = -0.5, size=3, col="blue")

other_interact = dplyr::select(data.frame(beta.sim),starts_with("raceother"))
other_fit<- cbind(lwr= apply(other_interact, 2, quant5)[1,], fit= apply(other_interact, 2, quant5)[2,], upr=apply(other_interact, 2, quant5)[3,])

ggplot(as.data.frame(other_fit), aes(rownames(other_fit), 
  fit, size=10, group=1, ylim=max(upr)+0.8)) +
  theme_bw(base_size=10)+ 
  geom_point(size=2)+
  geom_line(size=1)+
  geom_errorbar(aes(ymin = lwr, ymax = upr), width = 0.2, size=1)+
  xlab("Lab")+
  ylab("Fitted value")+
  ggtitle("Gestational Age vs. other interaction terms (JAGS)")+
  theme(axis.text.x=element_blank())+
  geom_text(aes(label = round(lwr,2), y = lwr), vjust = 1.5, size=3) +
  geom_text(aes(label = round(upr,2), y = upr), vjust = -0.5, size=3) +
  geom_text(aes(label = rownames(other_fit), y = upr+0.5), vjust = -0.5, size=3, col="blue")

### center 10 higher than center 5
### pcbs all cover 0, but 105, 170 higher, and 194 lower
### black lower than white; black:center10 higher and large variance, balck:center 31 higher, black:center 55 66 higher; black:center 60lower
### other:center 10 higher; other:center 37 lower; other:center 60 66 lower; 
```

Without all interaction terms

```{r}
full.lm<- lm(gestational_age ~ dde + pcb_028 + pcb_052 + pcb_074 + pcb_105 + pcb_118 + pcb_153+ pcb_170 + pcb_138 + pcb_180 + pcb_194 + pcb_203 + race + triglycerides + cholesterol+ maternal_age + smoking_status + center, data = data_age_na)

Y = data_age_na$gestational_age
X0 = model.matrix(full.lm)
# remove X where the fitted value is NA
coef(full.lm)[is.na(coef(full.lm))] #EE0.01:labBayer  EE0.03:labBayer EE0.01:labZeneca 
namesX0 = colnames(X0)
# X0 = X0[,-c(which(namesX0=="raceblack:center15"),
#            which(namesX0=="raceother:center15"),
#            which(namesX0=="raceother:center45"),
#            which(namesX0=="raceblack:center50"),
#            which(namesX0=="raceother:center82"))]
X1 = dplyr::select(data.frame(X0),starts_with("center"))
X2 = dplyr::select(data.frame(X0),-starts_with("center"))[,-1]
X = as.matrix(cbind(X1,X2))
model = function(){
  for (i in 1:n){
    Y[i] ~ dnorm(alpha + Xs[i,] %*% beta, phi)
  }
  
  alpha ~ dnorm(0, .000001*phi)
  phi ~ dgamma(.001, .001)
  sigma_L ~ dt(0,1,1)
  phi_L <- 1/((sigma_L^2)+.000001)
  tau ~ dgamma(.5, .5*n)
  sigma <- sqrt(1/phi)
  
  # beta for lab
  for (j in 1:p_center){
    lambda[j] ~ dgamma(a/2, a/2)
    beta[j] ~ dnorm(0, phi*phi_L*lambda[j]) # beta for lab starts from second column
  }
  
  # beta for others
  beta[(p_center+1):p] ~ dmnorm(rep(0,p-p_center), phi*tau*SSX2)
}
set.seed(1)
Xs = scale(X)
data = list(Y=Y, Xs=Xs, SSX2 = t(Xs[,(ncol(X1)+1):ncol(X)])%*%Xs[,(ncol(X1)+1):ncol(X)],
            n=length(Y), p_center=ncol(X1), p=ncol(X), a=9)
output = jags(data, inits=NULL,
              parameters.to.save=c("alpha","beta","sigma","sigma_L","lambda"),
              model=model, n.iter=10000)
sim.matrix = output$BUGSoutput$sims.matrix


n = length(Y)
S = diag(c(apply(X,2,function(x) {var(x)})),nrow=ncol(X))
# Transform back original betas
beta.sim = sim.matrix[,2:30] %*% (solve(S)^0.5) #beta_s = S^0.5*beta
#max(abs(beta.sim %*% (S^0.5) - sim.matrix[,2:57])) # Checkings
# Transform back original alpha
alphaS.sim = sim.matrix[,1]
Xbar = apply(X,2,mean)
alpha.sim = alphaS.sim - beta.sim %*% Xbar
colnames(alpha.sim) = "(Intercept)"
colnames(beta.sim) = colnames(X)

quant5 = function(x) {round(quantile(x, c(0.025, 0.5, 0.975)),2)} ## 95% CI
quant10 = function(x) {round(quantile(x, c(0.05, 0.5, 0.95)),2)} ## 90% CI
apply(beta.sim, 2, quant5)
apply(beta.sim, 2, quant10)

gelman.diag(as.mcmc(output))
### However, mixing is not good for sigma_L, which is the variation between centers. 
```


```{r}
full.lm<- lm(gestational_age ~ dde + pcb_028 + pcb_074 + pcb_105 + pcb_118 + pcb_153+ pcb_138 + pcb_180 + race + triglycerides + cholesterol+ maternal_age + smoking_status + center, data = data_age_na)

Y = data_age_na$gestational_age
X0 = model.matrix(full.lm)
# remove X where the fitted value is NA
coef(full.lm)[is.na(coef(full.lm))] #EE0.01:labBayer  EE0.03:labBayer EE0.01:labZeneca 
namesX0 = colnames(X0)
# X0 = X0[,-c(which(namesX0=="raceblack:center15"),
#            which(namesX0=="raceother:center15"),
#            which(namesX0=="raceother:center45"),
#            which(namesX0=="raceblack:center50"),
#            which(namesX0=="raceother:center82"))]
X1 = dplyr::select(data.frame(X0),starts_with("center"))
X2 = dplyr::select(data.frame(X0),-starts_with("center"))[,-1]
X = as.matrix(cbind(X1,X2))
model = function(){
  for (i in 1:n){
    Y[i] ~ dnorm(alpha + Xs[i,] %*% beta, phi)
  }
  
  alpha ~ dnorm(0, .000001*phi)
  phi ~ dgamma(.001, .001)
  sigma_L ~ dt(0,1,1)
  phi_L <- 1/((sigma_L^2)+.000001)
  tau ~ dgamma(.5, .5*n)
  sigma <- sqrt(1/phi)
  
  # beta for lab
  for (j in 1:p_center){
    lambda[j] ~ dgamma(a/2, a/2)
    beta[j] ~ dnorm(0, phi*phi_L*lambda[j]) # beta for lab starts from second column
  }
  
  # beta for others
  beta[(p_center+1):p] ~ dmnorm(rep(0,p-p_center), phi*tau*SSX2)
}
set.seed(1)
Xs = scale(X)
data = list(Y=Y, Xs=Xs, SSX2 = t(Xs[,(ncol(X1)+1):ncol(X)])%*%Xs[,(ncol(X1)+1):ncol(X)],
            n=length(Y), p_center=ncol(X1), p=ncol(X), a=9)
output = jags(data, inits=NULL,
              parameters.to.save=c("alpha","beta","sigma","sigma_L","lambda"),
              model=model, n.iter=10000)
sim.matrix = output$BUGSoutput$sims.matrix


n = length(Y)
S = diag(c(apply(X,2,function(x) {var(x)})),nrow=ncol(X))
# Transform back original betas
beta.sim = sim.matrix[,2:26] %*% (solve(S)^0.5) #beta_s = S^0.5*beta
#max(abs(beta.sim %*% (S^0.5) - sim.matrix[,2:57])) # Checkings
# Transform back original alpha
alphaS.sim = sim.matrix[,1]
Xbar = apply(X,2,mean)
alpha.sim = alphaS.sim - beta.sim %*% Xbar
colnames(alpha.sim) = "(Intercept)"
colnames(beta.sim) = colnames(X)

quant5 = function(x) {round(quantile(x, c(0.025, 0.5, 0.975)),2)} ## 95% CI
quant10 = function(x) {round(quantile(x, c(0.05, 0.5, 0.95)),2)} ## 90% CI
apply(beta.sim, 2, quant5)
apply(beta.sim, 2, quant10)

gelman.diag(as.mcmc(output))
### mixing is pretty good
```

```{r}
center = dplyr::select(data.frame(beta.sim),starts_with("center"))
center_fit<- cbind(lwr= apply(center, 2, quant5)[1,], fit= apply(center, 2, quant5)[2,], upr=apply(center, 2, quant5)[3,])

ggplot(as.data.frame(center_fit), aes(rownames(center_fit), 
  fit, size=10, group=1, ylim=max(upr)+0.8)) +
  theme_bw(base_size=10)+ 
  geom_point(size=2)+
  geom_line(size=1)+
  geom_errorbar(aes(ymin = lwr, ymax = upr), width = 0.2, size=1)+
  xlab("Lab")+
  ylab("Fitted value")+
  ggtitle("Gestational Age acorss centers with ref=5 (JAGS)")+
  theme(axis.text.x=element_blank())+
  geom_text(aes(label = round(lwr,2), y = lwr), vjust = 1.5, size=3) +
  geom_text(aes(label = round(upr,2), y = upr), vjust = -0.5, size=3) +
  geom_text(aes(label = rownames(center_fit), y = upr+0.5), vjust = -0.5, size=3, col="blue")


pcb = dplyr::select(data.frame(beta.sim),starts_with("pcb_"))
pcb_fit<- cbind(lwr= apply(pcb, 2, quant5)[1,], fit= apply(pcb, 2, quant5)[2,], upr=apply(pcb, 2, quant5)[3,])

ggplot(as.data.frame(pcb_fit), aes(rownames(pcb_fit), 
  fit, size=10, group=1, ylim=max(upr)+0.8)) +
  theme_bw(base_size=10)+ 
  geom_point(size=2)+
  geom_line(size=1)+
  geom_errorbar(aes(ymin = lwr, ymax = upr), width = 0.2, size=1)+
  xlab("Lab")+
  ylab("Fitted value")+
  ggtitle("Gestational Age acorss (JAGS)")+
  theme(axis.text.x=element_blank())+
  geom_text(aes(label = round(lwr,2), y = lwr), vjust = 1.5, size=3) +
  geom_text(aes(label = round(upr,2), y = upr), vjust = -0.5, size=3) +
  geom_text(aes(label = rownames(pcb_fit), y = upr+0.5), vjust = -0.5, size=3, col="blue")

# black_interact = dplyr::select(data.frame(beta.sim),starts_with("raceblack"))
# black_fit<- cbind(lwr= apply(black_interact, 2, quant5)[1,], fit= apply(black_interact, 2, quant5)[2,], upr=apply(black_interact, 2, quant5)[3,])

# ggplot(as.data.frame(black_fit), aes(rownames(black_fit), 
#   fit, size=10, group=1, ylim=max(upr)+0.8)) +
#   theme_bw(base_size=10)+ 
#   geom_point(size=2)+
#   geom_line(size=1)+
#   geom_errorbar(aes(ymin = lwr, ymax = upr), width = 0.2, size=1)+
#   xlab("Lab")+
#   ylab("Fitted value")+
#   ggtitle("Gestational Age vs. black interaction terms (JAGS)")+
#   theme(axis.text.x=element_blank())+
#   geom_text(aes(label = round(lwr,2), y = lwr), vjust = 1.5, size=3) +
#   geom_text(aes(label = round(upr,2), y = upr), vjust = -0.5, size=3) +
#   geom_text(aes(label = rownames(black_fit), y = upr+0.5), vjust = -0.5, size=3, col="blue")
# 
# other_interact = dplyr::select(data.frame(beta.sim),starts_with("raceother"))
# other_fit<- cbind(lwr= apply(other_interact, 2, quant5)[1,], fit= apply(other_interact, 2, quant5)[2,], upr=apply(other_interact, 2, quant5)[3,])
# 
# ggplot(as.data.frame(other_fit), aes(rownames(other_fit), 
#   fit, size=10, group=1, ylim=max(upr)+0.8)) +
#   theme_bw(base_size=10)+ 
#   geom_point(size=2)+
#   geom_line(size=1)+
#   geom_errorbar(aes(ymin = lwr, ymax = upr), width = 0.2, size=1)+
#   xlab("Lab")+
#   ylab("Fitted value")+
#   ggtitle("Gestational Age vs. other interaction terms (JAGS)")+
#   theme(axis.text.x=element_blank())+
#   geom_text(aes(label = round(lwr,2), y = lwr), vjust = 1.5, size=3) +
#   geom_text(aes(label = round(upr,2), y = upr), vjust = -0.5, size=3) +
#   geom_text(aes(label = rownames(other_fit), y = upr+0.5), vjust = -0.5, size=3, col="blue")

### center 10 31 45 higher than center 5; center 15 37 82 lower than center 5
### pcbs all cover 0, but 105, 028 higher, and 074 lower
```


```{r}
PC1_pcb <- prcomp(data_age_na[,grepl("pcb", colnames(data_age_na))])$x[,1] 
data_age_na$PC1_pcb<- PC1_pcb

full.lm<- lm(gestational_age ~ dde + PC1_pcb + race + triglycerides + cholesterol+ maternal_age + smoking_status + center, data = data_age_na)

Y = data_age_na$gestational_age
X0 = model.matrix(full.lm)
# remove X where the fitted value is NA
coef(full.lm)[is.na(coef(full.lm))] #EE0.01:labBayer  EE0.03:labBayer EE0.01:labZeneca 
namesX0 = colnames(X0)
# X0 = X0[,-c(which(namesX0=="raceblack:center15"),
#            which(namesX0=="raceother:center15"),
#            which(namesX0=="raceother:center45"),
#            which(namesX0=="raceblack:center50"),
#            which(namesX0=="raceother:center82"))]
X1 = dplyr::select(data.frame(X0),starts_with("center"))
X2 = dplyr::select(data.frame(X0),-starts_with("center"))[,-1]
X = as.matrix(cbind(X1,X2))
model = function(){
  for (i in 1:n){
    Y[i] ~ dnorm(alpha + Xs[i,] %*% beta, phi)
  }
  
  alpha ~ dnorm(0, .000001*phi)
  phi ~ dgamma(.001, .001)
  sigma_L ~ dt(0,1,1)
  phi_L <- 1/((sigma_L^2)+.000001)
  tau ~ dgamma(.5, .5*n)
  sigma <- sqrt(1/phi)
  
  # beta for lab
  for (j in 1:p_center){
    lambda[j] ~ dgamma(a/2, a/2)
    beta[j] ~ dnorm(0, phi*phi_L*lambda[j]) # beta for lab starts from second column
  }
  
  # beta for others
  beta[(p_center+1):p] ~ dmnorm(rep(0,p-p_center), phi*tau*SSX2)
}
set.seed(1)
Xs = scale(X)
data = list(Y=Y, Xs=Xs, SSX2 = t(Xs[,(ncol(X1)+1):ncol(X)])%*%Xs[,(ncol(X1)+1):ncol(X)],
            n=length(Y), p_center=ncol(X1), p=ncol(X), a=9)
output = jags(data, inits=NULL,
              parameters.to.save=c("alpha","beta","sigma","sigma_L","lambda"),
              model=model, n.iter=10000)
sim.matrix = output$BUGSoutput$sims.matrix


n = length(Y)
S = diag(c(apply(X,2,function(x) {var(x)})),nrow=ncol(X))
# Transform back original betas
beta.sim = sim.matrix[,2:20] %*% (solve(S)^0.5) #beta_s = S^0.5*beta
#max(abs(beta.sim %*% (S^0.5) - sim.matrix[,2:57])) # Checkings
# Transform back original alpha
alphaS.sim = sim.matrix[,1]
Xbar = apply(X,2,mean)
alpha.sim = alphaS.sim - beta.sim %*% Xbar
colnames(alpha.sim) = "(Intercept)"
colnames(beta.sim) = colnames(X)

quant5 = function(x) {round(quantile(x, c(0.025, 0.5, 0.975)),2)} ## 95% CI
quant10 = function(x) {round(quantile(x, c(0.05, 0.5, 0.95)),2)} ## 90% CI
apply(beta.sim, 2, quant5)
apply(beta.sim, 2, quant10)

gelman.diag(as.mcmc(output))
### mixing for sigma_L is pretty bad
```




### Without pcb with lots of 0s
```{r}
full.lm<- lm(gestational_age ~ dde + pcb_028 + pcb_074 + pcb_105 + pcb_118 + pcb_153+ pcb_138 + pcb_180 + race + triglycerides + cholesterol+ maternal_age + smoking_status + center + race*dde + race*triglycerides + race*cholesterol+ race*maternal_age + race*smoking_status + center*race, data = data_age_na)

Y = data_age_na$gestational_age
X0 = model.matrix(full.lm)
# remove X where the fitted value is NA
coef(full.lm)[is.na(coef(full.lm))] #EE0.01:labBayer  EE0.03:labBayer EE0.01:labZeneca 
namesX0 = colnames(X0)
X0 = X0[,-c(which(namesX0=="raceblack:center15"),
           which(namesX0=="raceother:center15"),
           which(namesX0=="raceother:center45"),
           which(namesX0=="raceblack:center50"),
           which(namesX0=="raceother:center82"))]
X1 = dplyr::select(data.frame(X0),starts_with("center"))
X2 = dplyr::select(data.frame(X0),-starts_with("center"))[,-1]
X = as.matrix(cbind(X1,X2))
model = function(){
  for (i in 1:n){
    Y[i] ~ dnorm(alpha + Xs[i,] %*% beta, phi)
  }
  
  alpha ~ dnorm(0, .000001*phi)
  phi ~ dgamma(.001, .001)
  sigma_L ~ dt(0,1,1)
  phi_L <- 1/((sigma_L^2)+.000001)
  tau ~ dgamma(.5, .5*n)
  sigma <- sqrt(1/phi)
  
  # beta for lab
  for (j in 1:p_center){
    lambda[j] ~ dgamma(a/2, a/2)
    beta[j] ~ dnorm(0, phi*phi_L*lambda[j]) # beta for lab starts from second column
  }
  
  # beta for others
  beta[(p_center+1):p] ~ dmnorm(rep(0,p-p_center), phi*tau*SSX2)
}
set.seed(1)
Xs = scale(X)
data = list(Y=Y, Xs=Xs, SSX2 = t(Xs[,(ncol(X1)+1):ncol(X)])%*%Xs[,(ncol(X1)+1):ncol(X)],
            n=length(Y), p_center=ncol(X1), p=ncol(X), a=9)
output = jags(data, inits=NULL,
              parameters.to.save=c("alpha","beta","sigma","sigma_L","lambda"),
              model=model, n.iter=5000)
sim.matrix = output$BUGSoutput$sims.matrix


n = length(Y)
S = diag(c(apply(X,2,function(x) {var(x)})),nrow=ncol(X))
# Transform back original betas
beta.sim = sim.matrix[,2:53] %*% (solve(S)^0.5) #beta_s = S^0.5*beta
#max(abs(beta.sim %*% (S^0.5) - sim.matrix[,2:57])) # Checkings
# Transform back original alpha
alphaS.sim = sim.matrix[,1]
Xbar = apply(X,2,mean)
alpha.sim = alphaS.sim - beta.sim %*% Xbar
colnames(alpha.sim) = "(Intercept)"
colnames(beta.sim) = colnames(X)

quant5 = function(x) {round(quantile(x, c(0.025, 0.5, 0.975)),2)} ## 95% CI
quant10 = function(x) {round(quantile(x, c(0.05, 0.5, 0.95)),2)} ## 90% CI
apply(beta.sim, 2, quant5)
apply(beta.sim, 2, quant10)


gelman.diag(as.mcmc(output))
#traceplot(output)
```

*Part III*  Use JAGS (or your own MCMC) to fit your full model, but use a hierarchical prior for the laboratory effects to determine if interlaboratory variation is greater than variation due to measurement error:
$$\beta_l \mid  \sigma^2, \sigma^2_L, \lambda_l \sim N(0, \sigma^2 \sigma^2_L/\lambda_l)$$
where $\sigma^2$ is the variance of the response, 
$$\sigma_L \sim t_1$$ for $\sigma_L > 0$ is a half-Cauchy prior on $\sigma_L$, and $$\lambda_l \sim G(a/2, a/2)$$ allows for heavier tails than the normal for the laboratory effects in case any labs are "outliers". Taking $a$ to be infinity is the normal prior, while $a = 1$ is a Cauchy prior conditional on $\sigma^2 \sigma^2_L$. Think about what might values of $a$ might be appropriate in this context and discuss how you select $a$. Include a graph that illustrates your findings for each question as well as tables with estimates of effects of interest and the $\sigma$ and $\sigma_L$ with credible intervals.


# Final report

```{r}
library(ggplot2)
library(dplyr)
library(MASS)
library(ggpubr)
library(monomvn)
Longnecker <- as.data.frame(readRDS("Longnecker.rds"))
head(Longnecker)
Longnecker$center<- as.factor(as.character(Longnecker$center))
Longnecker$center <- relevel(Longnecker$center, ref = "5")
Longnecker[Longnecker$gestational_age > 46, "gestational_age"] <- 46 # Truncate at age == 46
data = Longnecker
data_age_na<- data[-which(is.na(data$pcb_028)), ]
```


## BAS (All pcbs are excluded from the model)
### dde, race, triglycerides are included in the model using BAS

## Proposed model (No interaction, and pcbs without lots of 0s)

### We have tried a model with all main effects and interactions between race and all others
### We tried a model with all main effects
### We tried a model with main effects without pcbs with lots of 0s (proposed) mixing is good for sigma_L
### We tried a model which replace the pcbs with PC1


```{r}
full.lm<- lm(gestational_age ~ dde + pcb_028 + pcb_074 + pcb_105 + pcb_118 + pcb_153+ pcb_138 + pcb_180 + race + triglycerides + cholesterol+ maternal_age + smoking_status + center, data = data_age_na)

Y = data_age_na$gestational_age
X0 = model.matrix(full.lm)
# remove X where the fitted value is NA
coef(full.lm)[is.na(coef(full.lm))] 
namesX0 = colnames(X0)

X1 = dplyr::select(data.frame(X0),starts_with("center"))
X2 = dplyr::select(data.frame(X0),-starts_with("center"))[,-1]
X = as.matrix(cbind(X1,X2))
model = function(){
  for (i in 1:n){
    Y[i] ~ dnorm(alpha + Xs[i,] %*% beta, phi)
  }
  
  alpha ~ dnorm(0, .000001*phi)
  phi ~ dgamma(.001, .001)
  sigma_L ~ dt(0,1,1)
  phi_L <- 1/((sigma_L^2)+.000001)
  tau ~ dgamma(.5, .5*n)
  sigma <- sqrt(1/phi)
  
  # beta for lab
  for (j in 1:p_center){
    lambda[j] ~ dgamma(a/2, a/2)
    beta[j] ~ dnorm(0, phi*phi_L*lambda[j]) # beta for lab starts from second column
  }
  
  # beta for others
  beta[(p_center+1):p] ~ dmnorm(rep(0,p-p_center), phi*tau*SSX2)
}
set.seed(1)
Xs = scale(X)
data = list(Y=Y, Xs=Xs, SSX2 = t(Xs[,(ncol(X1)+1):ncol(X)])%*%Xs[,(ncol(X1)+1):ncol(X)],
            n=length(Y), p_center=ncol(X1), p=ncol(X), a=9)
output = jags(data, inits=NULL,
              parameters.to.save=c("alpha","beta","sigma","sigma_L","lambda"),
              model=model, n.iter=10000)
sim.matrix = output$BUGSoutput$sims.matrix


n = length(Y)
S = diag(c(apply(X,2,function(x) {var(x)})),nrow=ncol(X))
# Transform back original betas
beta.sim = sim.matrix[,2:26] %*% (solve(S)^0.5) #beta_s = S^0.5*beta
#max(abs(beta.sim %*% (S^0.5) - sim.matrix[,2:57])) # Checkings
# Transform back original alpha
alphaS.sim = sim.matrix[,1]
Xbar = apply(X,2,mean)
alpha.sim = alphaS.sim - beta.sim %*% Xbar
colnames(alpha.sim) = "(Intercept)"
colnames(beta.sim) = colnames(X)

quant5 = function(x) {round(quantile(x, c(0.025, 0.5, 0.975)),2)} ## 95% CI
quant10 = function(x) {round(quantile(x, c(0.05, 0.5, 0.95)),2)} ## 90% CI
apply(beta.sim, 2, quant5)
apply(beta.sim, 2, quant10)

gelman.diag(as.mcmc(output))
### mixing is pretty good
```

```{r}
center = dplyr::select(data.frame(beta.sim),starts_with("center"))
center_fit<- cbind(lwr= apply(center, 2, quant5)[1,], fit= apply(center, 2, quant5)[2,], upr=apply(center, 2, quant5)[3,])

ggplot(as.data.frame(center_fit), aes(rownames(center_fit), 
  fit, size=10, group=1, ylim=max(upr)+0.8)) +
  theme_bw(base_size=10)+ 
  geom_point(size=2)+
  geom_line(size=1)+
  geom_errorbar(aes(ymin = lwr, ymax = upr), width = 0.2, size=1)+
  xlab("Lab")+
  ylab("Fitted value")+
  ggtitle("Gestational Age acorss centers with ref=5 (JAGS)")+
  theme(axis.text.x=element_blank())+
  geom_text(aes(label = round(lwr,2), y = lwr), vjust = 1.5, size=3) +
  geom_text(aes(label = round(upr,2), y = upr), vjust = -0.5, size=3) +
  geom_text(aes(label = rownames(center_fit), y = upr+0.5), vjust = -0.5, size=3, col="blue")


pcb = dplyr::select(data.frame(beta.sim),starts_with("pcb_"))
pcb_fit<- cbind(lwr= apply(pcb, 2, quant5)[1,], fit= apply(pcb, 2, quant5)[2,], upr=apply(pcb, 2, quant5)[3,])

ggplot(as.data.frame(pcb_fit), aes(rownames(pcb_fit), 
  fit, size=10, group=1, ylim=max(upr)+0.8)) +
  theme_bw(base_size=10)+ 
  geom_point(size=2)+
  geom_line(size=1)+
  geom_errorbar(aes(ymin = lwr, ymax = upr), width = 0.2, size=1)+
  xlab("Lab")+
  ylab("Fitted value")+
  ggtitle("Gestational Age acorss (JAGS)")+
  theme(axis.text.x=element_blank())+
  geom_text(aes(label = round(lwr,2), y = lwr), vjust = 1.5, size=3) +
  geom_text(aes(label = round(upr,2), y = upr), vjust = -0.5, size=3) +
  geom_text(aes(label = rownames(pcb_fit), y = upr+0.5), vjust = -0.5, size=3, col="blue")

### center 10 31 45 higher than center 5; center 15 37 82 lower than center 5
### pcbs all cover 0, but 105, 028 higher, and 074 lower
```