---
title: ""
author: "Youngsoo Baek, YunranChen and Xiaojun Zheng"
date: "1/20/2020"
output: 
  ioslides_presentation:
    widescreen: true
    smaller: true
---

<style>
slides > slide {
  overflow-x: auto !important;
  overflow-y: auto !important;
}
</style>


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(arm)
library(corrplot)
library(naniar)
library(ggplot2)
library(dplyr)
library(tibble)
library(tidyr)
library(purrr)
library(lme4)
data=readRDS("Longnecker.rds")
data=data%>%dplyr::select(gestational_age,dde:cholesterol,center)
```

# EDA

## EDA: Missing Data

- Over 90% missing `albumin`, around 20% missing `score_*`, others less than 0.1%
- Drop the `albumin` and keep the complete cases

```{r missing, echo=FALSE, message=FALSE, warning=FALSE}
gg_miss_upset(data)
```

## EDA: Correlation Plot

- Weak correlation between covariates and `gestational age`
- Large correlation among covariates especially for PCBs

```{r corrplot, echo=FALSE}
data_=data%>%mutate(race=as.numeric(race))
M2=cor(as.matrix(na.omit(data_)))
corrplot.mixed(M2,number.cex=.5, upper="ellipse")
```

## EDA: Histogram for All Variables

- 0-inflation on some of PCBs
- Long left tail of `gestational age` (after truncated at 46)

```{r hist, echo=FALSE, message=FALSE, warning=FALSE}
data_46=data_#%>%mutate(y=gestational_age)
data_46[data_46$gestational_age>46,"gestational_age"]=46
data_46=data_46%>%select(-albumin)%>%mutate(race=factor(race),center=factor(center))

data_46 %>%
  keep(is.numeric) %>% 
  gather() %>% 
  ggplot(aes(value)) +
    facet_wrap(~ key, scales = "free") +
    geom_histogram()
```

## EDA: Heterogeneity across Centers

The demographic background of each center varies a lot, suggesting heterogeity across centers.
(favor fixed effect instead of random effect)

```{r}
ggplot(data_46,mapping = aes(x=as.factor(center),y=gestational_age,color=race))+geom_boxplot()+theme_bw()
ggplot(data_46,mapping = aes(x=as.factor(center),y=score_income,color=race))+geom_boxplot()+theme_bw()
```


## EDA

```{r EDA}
ggplot(data = data) +
   geom_mosaic(aes(x = product(race, center), fill=race), na.rm=TRUE)+
  labs(x = "Center ID", y = "Race", title='Race by center')
```

## Challenges

  * Noisy measurement
  
  * Multicollinearity
  
  * Modeling Heterogeneity between centers

## Regression estimates

```{r, echo=F, include=T}
# Boilerplate
Longnecker <- readRDS("./Longnecker.rds")
Longnecker <- Longnecker[,colnames(Longnecker)!="albumin"] # Throw away albumin
Longnecker$center <- as.factor(Longnecker$center) # Factorize center
Longnecker[Longnecker$gestational_age > 46, "gestational_age"] <- 46 # Truncate at age == 46
x <- model.matrix(gestational_age ~ . - center, data = Longnecker)
miss_indices <- as.integer(row.names(x))
Longnecker_complete <- Longnecker[miss_indices,] # Only those non-missing obs.
PC1_pcb <- prcomp(x[,grepl("pcb", colnames(x))])$x[,1]
Longnecker_complete_pc <- cbind.data.frame(
  Longnecker_complete[,!grepl("pcb", colnames(Longnecker_complete))],
  PC1 = PC1_pcb)
lm_with_pc <- lm(gestational_age ~ ., data = Longnecker_complete_pc)
Longnecker_complete_pc$early <- (Longnecker_complete_pc$gestational_age < 37)
llm_with_pc <- glm(early ~ . - gestational_age, family = binomial, data = Longnecker_complete_pc)
lm_coefs <- c( coef(lm_with_pc)[c("dde", "PC1")],
   "center5" = as.numeric(coef(lm_with_pc)[1]),
   coef(lm_with_pc)[1] + coef(lm_with_pc)[ grepl("center", names(coef(lm_with_pc))) ] )
lm_confint <- rbind(confint(lm_with_pc)[c("dde", "PC1"),],
                    "center5" = confint(lm_with_pc)[1,],
                    confint(lm_with_pc)[1,] + 
                      confint(lm_with_pc)[grepl("center", rownames(confint(lm_with_pc))),] )
llm_coefs <- c( coef(llm_with_pc)[c("dde", "PC1")],
   "center5" = as.numeric(coef(llm_with_pc)[1]),
   coef(llm_with_pc)[1] + coef(llm_with_pc)[ grepl("center", names(coef(llm_with_pc))) ] )
suppressMessages(llm_confint <- rbind(confint(llm_with_pc)[c("dde", "PC1"),],
                    "center5" = confint(llm_with_pc)[1,],
                    confint(llm_with_pc)[1,] + 
                      confint(llm_with_pc)[grepl("center", rownames(confint(llm_with_pc))),] ) )
lm_coefs <- cbind.data.frame(Mean = lm_coefs, lm_confint)
llm_coefs <- cbind.data.frame(Mean = llm_coefs, llm_confint)
```

95\% confidence interval for Gaussian regression:
```{r, echo=F, include=T}
round(lm_coefs[c("dde", "PC1"),], 3)
```

For logistic regression:
```{r, echo=F, include=T}
round(llm_coefs[c("dde", "PC1"),], 3)
```

  * Hard to see improvement from a "baseline" model (Gaussian)

## Diagnostic Plots

```{r, echo=F, include=T}
par(mfrow = c(1, 2))
plot(lm_with_pc, 1, main = "Gaussian regression")
binnedplot(fitted(llm_with_pc),
           residuals(llm_with_pc, type = "response"),
           xlab = "Fitted response",
           ylab = "Residual",
           main = "Logistic regression (43 bins)")
```

```{r}
full.lm<- lm(gestational_age ~ dde + pcb_028 + pcb_074 + pcb_105 + pcb_118 + pcb_153+ pcb_138 + pcb_180 + race + triglycerides + cholesterol+ maternal_age + smoking_status + center, data = data_age_na)

Y = data_age_na$gestational_age
X0 = model.matrix(full.lm)
# remove X where the fitted value is NA
coef(full.lm)[is.na(coef(full.lm))] 
namesX0 = colnames(X0)

X1 = dplyr::select(data.frame(X0),starts_with("center"))
X2 = dplyr::select(data.frame(X0),-starts_with("center"))[,-1]
X = as.matrix(cbind(X1,X2))
model = function(){
  for (i in 1:n){
    Y[i] ~ dnorm(alpha + Xs[i,] %*% beta, phi)
  }
  
  alpha ~ dnorm(0, .000001*phi)
  phi ~ dgamma(.001, .001)
  sigma_L ~ dt(0,1,1)
  phi_L <- 1/((sigma_L^2)+.000001)
  tau ~ dgamma(.5, .5*n)
  sigma <- sqrt(1/phi)
  
  # beta for lab
  for (j in 1:p_center){
    lambda[j] ~ dgamma(a/2, a/2)
    beta[j] ~ dnorm(0, phi*phi_L*lambda[j]) # beta for lab starts from second column
  }
  
  # beta for others
  beta[(p_center+1):p] ~ dmnorm(rep(0,p-p_center), phi*tau*SSX2)
}
set.seed(1)
Xs = scale(X)
data = list(Y=Y, Xs=Xs, SSX2 = t(Xs[,(ncol(X1)+1):ncol(X)])%*%Xs[,(ncol(X1)+1):ncol(X)],
            n=length(Y), p_center=ncol(X1), p=ncol(X), a=9)
output = jags(data, inits=NULL,
              parameters.to.save=c("alpha","beta","sigma","sigma_L","lambda"),
              model=model, n.iter=10000)
sim.matrix = output$BUGSoutput$sims.matrix


n = length(Y)
S = diag(c(apply(X,2,function(x) {var(x)})),nrow=ncol(X))
# Transform back original betas
beta.sim = sim.matrix[,2:26] %*% (solve(S)^0.5) #beta_s = S^0.5*beta
#max(abs(beta.sim %*% (S^0.5) - sim.matrix[,2:57])) # Checkings
# Transform back original alpha
alphaS.sim = sim.matrix[,1]
Xbar = apply(X,2,mean)
alpha.sim = alphaS.sim - beta.sim %*% Xbar
colnames(alpha.sim) = "(Intercept)"
colnames(beta.sim) = colnames(X)

quant5 = function(x) {round(quantile(x, c(0.025, 0.5, 0.975)),2)} ## 95% CI
quant10 = function(x) {round(quantile(x, c(0.05, 0.5, 0.95)),2)} ## 90% CI
apply(beta.sim, 2, quant5)
apply(beta.sim, 2, quant10)

gelman.diag(as.mcmc(output))
### mixing is pretty good
```

```{r}
center = dplyr::select(data.frame(beta.sim),starts_with("center"))
center_fit<- cbind(lwr= apply(center, 2, quant5)[1,], fit= apply(center, 2, quant5)[2,], upr=apply(center, 2, quant5)[3,])

ggplot(as.data.frame(center_fit), aes(rownames(center_fit), 
  fit, size=10, group=1, ylim=max(upr)+0.8)) +
  theme_bw(base_size=10)+ 
  geom_point(size=2)+
  geom_line(size=1)+
  geom_errorbar(aes(ymin = lwr, ymax = upr), width = 0.2, size=1)+
  xlab("Lab")+
  ylab("Fitted value")+
  ggtitle("Gestational Age acorss centers with ref=5 (JAGS)")+
  theme(axis.text.x=element_blank())+
  geom_text(aes(label = round(lwr,2), y = lwr), vjust = 1.5, size=3) +
  geom_text(aes(label = round(upr,2), y = upr), vjust = -0.5, size=3) +
  geom_text(aes(label = rownames(center_fit), y = upr+0.5), vjust = -0.5, size=3, col="blue")


pcb = dplyr::select(data.frame(beta.sim),starts_with("pcb_"))
pcb_fit<- cbind(lwr= apply(pcb, 2, quant5)[1,], fit= apply(pcb, 2, quant5)[2,], upr=apply(pcb, 2, quant5)[3,])

ggplot(as.data.frame(pcb_fit), aes(rownames(pcb_fit), 
  fit, size=10, group=1, ylim=max(upr)+0.8)) +
  theme_bw(base_size=10)+ 
  geom_point(size=2)+
  geom_line(size=1)+
  geom_errorbar(aes(ymin = lwr, ymax = upr), width = 0.2, size=1)+
  xlab("Lab")+
  ylab("Fitted value")+
  ggtitle("Gestational Age acorss (JAGS)")+
  theme(axis.text.x=element_blank())+
  geom_text(aes(label = round(lwr,2), y = lwr), vjust = 1.5, size=3) +
  geom_text(aes(label = round(upr,2), y = upr), vjust = -0.5, size=3) +
  geom_text(aes(label = rownames(pcb_fit), y = upr+0.5), vjust = -0.5, size=3, col="blue")

### center 10 31 45 higher than center 5; center 15 37 82 lower than center 5
### pcbs all cover 0, but 105, 028 higher, and 074 lower
```


## Result

## Conclusion
