---
title: "Effect of DDE and PCB Exposure on Pre-Term Delivery"
author: "Youngsoo Baek, Yunran Chen, Xiaojun Zheng"
output: beamer_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.height = 5)
library(arm)
library(corrplot)
library(naniar)
library(ggplot2)
library(dplyr)
library(tibble)
library(tidyr)
library(purrr)
library(lme4)
library(MASS)
library(ggmosaic)
```

## Missing data 

- Over 90% missing `albumin`, around 20% missing `score_*`, others less than 0.1%
- Drop the `albumin` and keep the complete cases

```{r missing, echo=FALSE, message=FALSE, warning=FALSE,cache=TRUE}
data=readRDS("Longnecker.rds")
data=data %>% dplyr::select(gestational_age, dde:cholesterol, center)
gg_miss_upset(data)
```

## Noisy measurement

- Zero-inflation on some of PCBs -> PCA vs Variable Selection
- Long left tail of `gestational age` -> categorical variable

```{r hist, echo=FALSE, message=FALSE, warning=FALSE,cache=TRUE}
data_46=data_#%>%mutate(y=gestational_age)
data_46[data_46$gestational_age>46,"gestational_age"]=46
data_46=data_46%>%select(-albumin)%>%mutate(race=factor(race),center=factor(center))

data_46$earlylvl <- cut(data_46$gestational_age,
                                       breaks = c(0, 31, 33, 36, 46), right = TRUE)
data_46 %>%
  dplyr::select(gestational_age:triglycerides,-triglycerides,-dde) %>%
  gather() %>% 
  ggplot(aes(value)) +
    facet_wrap(~ key, scales = "free") +
    geom_histogram() + theme_bw()
```


## Multicollinearity

- Weak correlation between covariates and `gestational age` (less than 0.3)
- Large correlation among covariates especially for PCBs

```{r corrplot, echo=FALSE,cache=TRUE}
data_ = data %>% mutate(race=as.numeric(race))
M2 = cor(as.matrix(na.omit(data_%>%select(pcb_028:pcb_203))))
corrplot.mixed(M2,number.cex=2, upper="ellipse")
#M3 = cor(as.matrix(na.omit(data_)))
#M3[1,-1]
#cordf=data.frame(cor=M3[1,-1],var=names(M3[1,-1]))
#ggplot(cordf,aes(x=var,y=cor))+geom_point()+theme_bw()
```

## Heterogeneity across Centers: fix effects

The demographic background (race, income, etc.) vary across center. -> fix effects

```{r heter, echo=FALSE, cache=TRUE}
ggplot(data%>%filter(center%in%c(5,10,15,31,45,60)),mapping = aes(x=as.factor(center),y=gestational_age,color=race))+geom_boxplot()+theme_bw()
```

## Ultimate goal of the model
What would a hypothetical experimental study for DDE and PCB's look like?

$$
\begin{aligned}
\text{Null model:}\quad\text{gestational age} \sim 1 + \text{Demographic variables} \\
\text{Alternative:}\quad\text{gestational age} \sim 1 + DDE + PCB + \text{...} + \text{Demographic}
\end{aligned}
$$

## Addressing PCB's
Two main approaches:

  i. Everyone contributes, with different weights (**principal component regression**)
  
  ii. Pick a few representative voters (**variable selection**)

```{r boilerplate_for_models, echo=F, include=F}
Longnecker <- readRDS("./Longnecker.rds")
Longnecker <- Longnecker[,colnames(Longnecker)!="albumin"] # Throw away albumin
Longnecker$center <- as.factor(Longnecker$center) # Factorize center
Longnecker[Longnecker$gestational_age > 46, "gestational_age"] <- 46 # Truncate at age == 46
complete_formula <- as.formula(
  "gestational_age ~ dde + pcb_028 + pcb_052 + pcb_074 + pcb_105 + 
  pcb_118 + pcb_153 + pcb_170 + pcb_138 + pcb_180 + pcb_194 + pcb_203 + 
  triglycerides + race + maternal_age + smoking_status + cholesterol + center"
)
x <- model.matrix(complete_formula, data = Longnecker)
miss_indices <- as.integer(row.names(x))
Longnecker_complete <- Longnecker[miss_indices,] # Only those non-missing obs.
PC1_pcb <- prcomp(x[,grepl("pcb", colnames(x))], scale.=TRUE)$x[,1]
Longnecker_complete_pc <- cbind.data.frame(
  Longnecker_complete[,!grepl("pcb", colnames(Longnecker_complete))],
  PC1 = PC1_pcb)
# Mutation: gestational week => categorical variables
Longnecker_complete_pc$early <- (Longnecker_complete_pc$gestational_age < 37)
Longnecker_complete_pc$earlylvl <- cut(Longnecker_complete_pc$gestational_age,
                                       breaks = c(0, 31, 33, 36, 46), right = TRUE)
```

## Model: Logistic vs. Ordinal logistic

  - Binary response: preterm delivery (<37 wks) 
  
    - Loss of information about different levels of risk involved in ordered levels

```{r, echo=F, include=T}
Longnecker_complete_pc$earlylvl %>% 
  table(.) %>% prop.table() %>% 
  barplot(names.arg = c("Very preterm\n(<32)", "Moderate\n(32-36)",
                        "Late preterm\n(<36)", "Normal--Late\n(>=37)"))
```

    
## Interpreting the model

  - Logistic: coefficients $\beta$ correspond to $\times e^\beta$ increase in the preterm delivery *odds*
  
  - Ordinal logistic: Assumes multiple (>2) delivery category odds are *proportional*
  
$${\rm Pr}(Y_i \leq k|X_i) = {\rm logit}^{-1}(\beta_{0,k} + \boldsymbol{\beta}^TX_i),\; k=1,2,3,4.$$
  
  - Possible violation: can be proportional, but not by a constant factor

## Predictors to be adjusted for

  - Excluded: three score variables relating to education, income, and occupation
  
  - Justification: $F$-test against other predictors excluding chemicals, exploratory model fits
    
  - First principal component for PCB levels (scaled)

```{r, echo=F, include=T}
pca_formula1 <- as.formula("early ~ dde + PC1 + 
  triglycerides + race + maternal_age + smoking_status + cholesterol + center")
pca_formula2 <- as.formula("earlylvl ~ dde + PC1 + 
  triglycerides + race + maternal_age + smoking_status + cholesterol + center")
# 1. Binary logistic
llm_with_pc <- glm(pca_formula1, family = binomial, 
                   data = Longnecker_complete_pc)
# 2. Ordinal logistic
olm_with_pc <- polr(pca_formula2, data = Longnecker_complete_pc)
#
llm_coefs <- coef(llm_with_pc)
llm_confint <- suppressMessages(confint(llm_with_pc))

llm_coefs <- cbind.data.frame(Mean = llm_coefs, llm_confint)
#
olm_coefs <- -1 * coef(olm_with_pc)
olm_confint <- suppressMessages(-1 * confint(olm_with_pc)[,c(2,1)])
olm_coefs <- suppressMessages(cbind.data.frame(Mean = olm_coefs, olm_confint))
colnames(olm_coefs) <- colnames(olm_coefs)[c(1,3,2)]
```



## Estimated Effects
 
 - 95\% confidence interval estimates for significant coefficients
 
```{r, echo=F, include=T}
knitr::kable(llm_coefs[c("dde", "PC1", "triglycerides", "cholesterol"),], format = "latex", digits = 3,
             caption = "Logistic")
knitr::kable(olm_coefs[c("dde", "PC1", "triglycerides"),], format = "latex", digits = 3,
             caption = "Ordinal logistic")
```

## Estimated Effects

  - Two models agree in significant positive mean shifts for center IDs 15, 37, 82 (large number of black subjects)

 - "Baseline" log odds +/- 2 standard errors, on probability scale, is estimated for each category by the ordered logistic model.

```{r, echo=F, message=F, include=T}
blo <- summary(olm_with_pc)$coefficients[20:22, "Value"] # baseline log-odds
bprobs <- exp(blo) / (1+exp(blo))
# Transforming +/- 2SE to probability scale
odds_lb <- exp(blo) * exp(-2 * summary(olm_with_pc)$coefficients[20:22, "Std. Error"])
odds_ub <- exp(blo) * exp(2 * summary(olm_with_pc)$coefficients[20:22, "Std. Error"])
probs_lb <- odds_lb / (1 + odds_lb); probs_ub <- odds_ub / (1 + odds_ub) 
knitr::kable(cbind.data.frame("Lower bound" = probs_lb,
                              "Mean" = bprobs,
                              "Upper bound" = probs_ub,
                              row.names = c("Very preterm", "Moderately", "Late preterm")
                              ),
             format = "latex", digits = 3)
```


## Interpretation

  - Model 1: Adjusted for PCB levels and demographic variables, a 1ug increase in DDE exposure corresponds to `r round(exp(llm_coefs[c("dde"), 1]),3)` times more odds of preterm delivery.
  
  - Model 2: (Adjusted) A 1ug increase in DDE exposure corresponds to `r round(exp(olm_coefs[c("dde"), 1]),3)` times more odds of more severely preterm delivery (very than moderately so, etc.).

  - Similar interpretation can be done for `PC1` and individual weights given to PCB compounds, since the weights are all positive
  
  - However, inference is unidentical to DDE in the sense that we are not adjusting for other PCB compoounds

## Diagnostic Plots

```{r, echo=F, include=T}
par(mfrow = c(1, 2))
binnedplot(fitted(llm_with_pc),
           residuals(llm_with_pc, type = "response"),
           nclass = 40,
           xlab = "Fitted response",
           ylab = "Residual",
           main = "Logistic regression (40 bins)")
est_probs_llm <- predict(llm_with_pc, type = "response")
boxplot(est_probs_llm ~ Longnecker_complete_pc$early,
        xlab = "Preterm delivery",
        ylab = "Estimated probabilities")
#
est_probs_olm <- predict(olm_with_pc, type = "probs")
```

## Further Discussion: PCBs

- Aggregating information of all the PCB's (PCA -> hard to interpret)
- Selecting representative PCB's (Frequentist and Bayesian variable selection)
  - Bayesian Model Averaging 
  - Bayesian Lasso (PCB 074, 153)
  - Horseshoe Prior
  - Lasso (Logistic: dde, pcb_074, 153) -> 'consistent' with the previous model

```{r}
load("./lasso_logistic.rdata")
lasso_fit<- lasso_logistic$fit
a<- data.frame(name = colnames[1:3], beta = lasso_fit[c(1:3),3])

knitr::kable(a, format = "latex", digits = 3,
             caption = "Logistic")
```


```{r blasso}
load("./lasso_logistic.rdata")
lasso_fit<- lasso_logistic$fit
a<- data.frame(name = colnames[1:3], beta = lasso_fit[c(1:3),3])
colnames<- colnames(X)[lasso_logistic$fit[,1]+1]


new<- as.data.frame(cbind(data_age_na[, colnames(data_age_na) %in% colnames], data_age_na$gestational_age_binary))
#predict(lasso_logistic, new, interval = "prediction")

yhat_G <- predict(lasso_logistic, new, bayes.avg=TRUE)
lines(df$X, yhat_G[,1], col="blue", lwd=2.5)
lines(df$X, yhat_G[,3], col="blue", lwd=1, lty="dashed")
lines(df$X, yhat_G[,4], col="blue", lwd=1, lty="dashed")


legend(1,11,c("Gaussian","Student-t (dof=5)"),lty=c(1,1),col=c("blue","darkred"),
       lwd=c(2.5,2.5), cex=0.7)


colnames(beta.sim) = colnames(X)[-1]
quant5 = function(x) {round(quantile(x, c(0.025, 0.5, 0.975)),2)} ## 95% CI
pcb = dplyr::select(data.frame(beta.sim),starts_with("pcb_"))
pcb_fit<- cbind(lwr= apply(pcb, 2, quant5)[1,], fit= apply(pcb, 2, quant5)[2,], upr=apply(pcb, 2, quant5)[3,])

ggplot(as.data.frame(pcb_fit), aes(rownames(pcb_fit), 
  fit, size=10, group=1, ylim=max(upr)+0.8)) +
  geom_point(size=2)+
  geom_line(size=1)+
  geom_errorbar(aes(ymin = lwr, ymax = upr), width = 0.2, size=1)+
  xlab("Lab")+
  ylab("Fitted value")+
  ggtitle("Gestational Age acorss (Lasso)")+
  theme(axis.text.x=element_blank())+
  geom_text(aes(label = round(lwr,2), y = lwr), vjust = 1.5, size=3) +
  geom_text(aes(label = round(upr,2), y = upr), vjust = -0.5, size=3) +
  geom_text(aes(label = rownames(pcb_fit), y = upr+0.5), vjust = -0.5, size=3, col="blue")

exp(quant5(beta.sim[,1]))-1
```
 
```{r blasso0}
load("./beta.sim_blasso0.rdata")
colnames(beta.sim) = colnames(X)[-1]
quant5 = function(x) {round(quantile(x, c(0.025, 0.5, 0.975)),2)} ## 95% CI
pcb = dplyr::select(data.frame(beta.sim),starts_with("pcb_"))
pcb_fit<- cbind(lwr= apply(pcb, 2, quant5)[1,], fit= apply(pcb, 2, quant5)[2,], upr=apply(pcb, 2, quant5)[3,])

ggplot(as.data.frame(pcb_fit), aes(rownames(pcb_fit), 
  fit, size=10, group=1, ylim=max(upr)+0.8)) +
  geom_point(size=2)+
  geom_line(size=1)+
  geom_errorbar(aes(ymin = lwr, ymax = upr), width = 0.2, size=1)+
  xlab("Lab")+
  ylab("Fitted value")+
  ggtitle("Gestational Age acorss (Lasso)")+
  theme(axis.text.x=element_blank())+
  geom_text(aes(label = round(lwr,2), y = lwr), vjust = 1.5, size=3) +
  geom_text(aes(label = round(upr,2), y = upr), vjust = -0.5, size=3) +
  geom_text(aes(label = rownames(pcb_fit), y = upr+0.5), vjust = -0.5, size=3, col="blue")

exp(quant5(beta.sim[,1]))-1
```

```{r, cache=T,echo=F,eval=F}
# switch to T to include; I don't have your file!!
# load("./HP_pcb.rdata")
# quant5 = function(x) {round(quantile(x, c(0.025, 0.5, 0.975)),2)} ## 95% CI
# pcb = dplyr::select(data.frame(beta.sim),starts_with("pcb_"))
# pcb_fit<- cbind(lwr= apply(pcb, 2, quant5)[1,], fit= apply(pcb, 2, quant5)[2,], upr=apply(pcb, 2, quant5)[3,])
# 
# ggplot(as.data.frame(pcb_fit), aes(rownames(pcb_fit), 
#   fit, size=10, group=1, ylim=max(upr)+0.8)) +
#   geom_point(size=2)+
#   geom_line(size=1)+
#   geom_errorbar(aes(ymin = lwr, ymax = upr), width = 0.2, size=1)+
#   xlab("Lab")+
#   ylab("Fitted value")+
#   ggtitle("Gestational Age acorss (JAGS)")+
#   theme(axis.text.x=element_blank())+
#   geom_text(aes(label = round(lwr,2), y = lwr), vjust = 1.5, size=3) +
#   geom_text(aes(label = round(upr,2), y = upr), vjust = -0.5, size=3) +
#   geom_text(aes(label = rownames(pcb_fit), y = upr+0.5), vjust = -0.5, size=3, col="blue")
```


```{r lasso, cache=T,echo=F, eval=T}
library("glmnet")
data=readRDS("Longnecker.rds")
data=data%>%dplyr::select(gestational_age,dde:cholesterol,center)
data_46=data%>%mutate(y=gestational_age)
data_46[data_46$y>46,"y"]=46
data_46=data_46%>%dplyr::select(-gestational_age,-albumin)%>%mutate(race=factor(race),center=factor(center))
data_46_wscore=data_46%>%na.omit()
set.seed(123)

train.data=data_46_wscore
x <- model.matrix(y~., train.data)[,-1]
# Convert the outcome (class) to a numerical variable
y <- ifelse(train.data$y<37,1,0)
train.data$y=y
cv.lasso2 <- cv.glmnet(x, y, alpha = 1,family="binomial")
# Fit the final model on the training data
model2 <- glmnet(x, y, alpha = 1,family="binomial",
                lambda = cv.lasso2$lambda[27])

plot(cv.lasso2) # logistic
cv.lasso2$lambda
cv.lasso2$lambda.min
which(cv.lasso2$lambda==cv.lasso2$lambda.min)
coef(model2) # dde, pcb_074, 153; the one with min mse, if relax the penalty a little bit, will include all the demographic characters
```

```{r glmlasso, cache=T,echo=F, eval=T}
formula_lasso_w <- as.formula(
  "y ~ dde + pcb_074 + pcb_153 + 
  triglycerides + race + maternal_age + smoking_status + cholesterol + center + score_education + score_income + score_occupation"
)
formula_lasso_o <- as.formula(
  "y ~ dde + 
  triglycerides + race + maternal_age + smoking_status + cholesterol + center + score_education + score_income + score_occupation"
)
model21_w <- glm(formula_lasso_w, family = binomial, 
                   data = train.data)
summary(model21_w)
model21_o <- glm(formula_lasso_o, family = binomial, 
                   data = train.data)
anova(model21_o,model21_w,test="LRT")
```

## inference on lasso

```{r}
library("selectiveInference")
model2min <- glmnet(x, y, alpha = 1,family="binomial",
                lambda = cv.lasso2$lambda.min)

beta=coef(model2min)
X=model.matrix(y~.,data=train.data)
beta=as.matrix(beta)
cis=fixedLassoInf(x=X[,-1], y=(train.data$y), beta=beta, lambda=cv.lasso2$lambda.min,family = "binomial",alpha=0.05)
cis
```

## ordinal logistic regression.

```{r vs, cache=T,echo=F}
library("ordinalNet")
train.data.ord=data_46_wscore
train.data.ord$y <- cut(train.data.ord$y,
                                       breaks = c(0, 31, 33, 36, 46), right = TRUE)
ord.x=model.matrix(y~.,data=train.data.ord)
ordlasso.model <- ordinalNet(ord.x, train.data.ord$y, family="cumulative", link="logit",
                   parallelTerms=TRUE, nonparallelTerms=FALSE,alpha=1)
coef(ordlasso.model) # pcb_028,074,153,138


formula_lasso_w.ord <- as.formula(
  "y ~ dde + pcb_074 + pcb_153 +
  triglycerides + race + maternal_age + smoking_status + cholesterol + center + score_education + score_income + score_occupation"
)
formula_lasso_o.ord <- as.formula(
  "y ~ dde + 
  triglycerides + race + maternal_age + smoking_status + cholesterol + center + score_education + score_income + score_occupation"
)

model21_w.ord <- polr(formula_lasso_w.ord, 
                   data = train.data.ord)
summary(model21_w.ord)
model21_o.ord <- polr(formula_lasso_o.ord, 
                   data = train.data.ord)
summary(model21_o.ord)
anova(model21_o.ord,model21_w.ord)
```

## Possible Improvements

  - Pooling heterogeneous effects across centers
  
  - Incorporating interactions: systematic, priors-based approach
  
  - Different methods to tackle nonlinearity
  
