---
title: "Effect of DDE and PCB Exposure on Pre-Term Delivery"
author: "Youngsoo Baek, Yunran Chen, Xiaojun Zheng"
output: beamer_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(arm)
library(corrplot)
library(naniar)
library(ggplot2)
library(dplyr)
library(tibble)
library(tidyr)
library(purrr)
library(lme4)
library(MASS)
#library(ggpubr)
#library(monomvn)
#library(BAS)
#library(R2jags)
library(ggmosaic)
```

## EDA for missing data

- Over 90% missing `albumin`, around 20% missing `score_*`, others less than 0.1%
- Drop the `albumin` and keep the complete cases

```{r missing, echo=FALSE, message=FALSE, warning=FALSE,cache=TRUE}
data=readRDS("Longnecker.rds")
data=data %>% dplyr::select(gestational_age, dde:cholesterol, center)
gg_miss_upset(data)
```

## EDA: Correlation Plot

- Weak correlation between covariates and `gestational age`
- Large correlation among covariates especially for PCBs

```{r corrplot, echo=FALSE,cache=TRUE}
data_ = data %>% mutate(race=as.numeric(race))
M2 = cor(as.matrix(na.omit(data_)))
corrplot.mixed(M2,number.cex=.5, upper="ellipse")
```

## EDA: Histogram for All Variables

- Zero-inflation on some of PCBs
- Long left tail of `gestational age` (after truncated at 46)

```{r hist, echo=FALSE, message=FALSE, warning=FALSE,cache=TRUE}
data_46=data_#%>%mutate(y=gestational_age)
data_46[data_46$gestational_age>46,"gestational_age"]=46
data_46=data_46%>%select(-albumin)%>%mutate(race=factor(race),center=factor(center))

data_46 %>%
  keep(is.numeric) %>% 
  gather() %>% 
  ggplot(aes(value)) +
    facet_wrap(~ key, scales = "free") +
    geom_histogram()
```

## EDA: Heterogeneity across Centers

The demographic background of each center varies a lot, suggesting heterogeity across centers.
(favor fixed effect instead of random effect)

```{r heter, echo=FALSE, cache=TRUE}
ggplot(data_46,mapping = aes(x=as.factor(center),y=gestational_age,color=race))+geom_boxplot()+theme_bw()
ggplot(data_46,mapping = aes(x=as.factor(center),y=score_income,color=race))+geom_boxplot()+theme_bw()
```

```{r EDA, echo=FALSE, cache=TRUE}
Longnecker <- readRDS("./Longnecker.rds")
Longnecker <- Longnecker[,colnames(Longnecker)!="albumin"] # Throw away albumin
Longnecker$center <- as.factor(Longnecker$center) # Factorize center
Longnecker[Longnecker$gestational_age > 46, "gestational_age"] <- 46 # Truncate at age == 46
data = Longnecker
data_age_na<- data[-which(is.na(data$pcb_028)), ]

ggplot(data = data_age_na) +
   geom_mosaic(aes(x = product(race, center), fill=race), na.rm=TRUE)+
  labs(x = "Center ID", y = "Race", title='Race by center') + theme_bw()
```

## Challenges

  * Noisy measurement
  
  * Multicollinearity
  
  * Modeling Heterogeneity between centers

## Ultimate goal of the model
What would a hypothetical experimental study for DDE and PCB's look like?

$$
\begin{aligned}
\text{Null model:}\quad\text{gestational age} \sim 1 + \text{Demographic variables} \\
\text{Alternative:}\quad\text{gestational age} \sim 1 + DDE + PCB + \text{...} + \text{Demographic}
\end{aligned}
$$

```{r boilerplate_for_models, echo=F, include=F}
Longnecker <- readRDS("./Longnecker.rds")
Longnecker <- Longnecker[,colnames(Longnecker)!="albumin"] # Throw away albumin
Longnecker$center <- as.factor(Longnecker$center) # Factorize center
Longnecker[Longnecker$gestational_age > 46, "gestational_age"] <- 46 # Truncate at age == 46
x <- model.matrix(gestational_age ~ . - center, data = Longnecker)
miss_indices <- as.integer(row.names(x))
Longnecker_complete <- Longnecker[miss_indices,] # Only those non-missing obs.
PC1_pcb <- prcomp(x[,grepl("pcb", colnames(x))], scale.=TRUE)$x[,1]
Longnecker_complete_pc <- cbind.data.frame(
  Longnecker_complete[,!grepl("pcb", colnames(Longnecker_complete))],
  PC1 = PC1_pcb)
# Mutation: gestational week => categorical variables
Longnecker_complete_pc$early <- (Longnecker_complete_pc$gestational_age < 37)
Longnecker_complete_pc$earlylvl <- cut(Longnecker_complete_pc$gestational_age,
                                       breaks = c(0, 31, 33, 36, 46), right = TRUE)
```

## Model: Logistic vs. Ordinal logistic

  - Binary response: preterm delivery (<37 wks) 
  
    - Loss of information about different levels of risk involved in different periods
    
  - Ordinal response: categories of delivery periods
  
    i. Very preterm (< 32 wks)
    
    ii. Moderately preterm (32 or 33 wks)
    
    iii. Late preterm (majority; < 36 wks)

```{r, echo=F, include=T}
Longnecker_complete_pc$earlylvl %>% 
  table(.) %>% prop.table() %>% 
  barplot()
```

    
## Interpreting the model

  - Logistic: coefficients $\beta$ correspond to $\times e^\beta$ increase in the preterm delivery *odds*
  
  - Ordinal logistic: Assumes multiple (>2) delivery category odds are *proportional*
  
$${\rm Pr}(Y_i \leq k|X_i) = {\rm logit}^{-1}(\beta_{0,k} + \boldsymbol{\beta}^TX_i),\; k=1,2,3,4.$$
  
    - Possible violation: can be proportional, but not by a constant factor

## Predictors to be adjusted for

  - Indication from model fits against including score variables
  
    - Advantage of resolving missingness issues
    
  - First principal component for PCB levels (scaled)
  
    - Multicollinearity abetted at cost of direct interpretation of effect size

(... Want some kind of plot here illustrating the above points ...)


```{r, echo=F, include=T}
# 1. Binary logistic
llm_with_pc <- glm(early ~ . - earlylvl - gestational_age, family = binomial, 
                   data = Longnecker_complete_pc)
# 2. Ordinal logistic
olm_with_pc <- polr(earlylvl ~ . - early - gestational_age, data = Longnecker_complete_pc)
#
llm_coefs <- c( coef(llm_with_pc)[c("dde", "PC1")],
   "center5" = as.numeric(coef(llm_with_pc)[1]),
   coef(llm_with_pc)[1] + coef(llm_with_pc)[ grepl("center", names(coef(llm_with_pc))) ] )
llm_confint <- suppressMessages(confint(llm_with_pc))
llm_confint <- suppressMessages(rbind(llm_confint[c("dde", "PC1"),],
                    "center5" = llm_confint[1,],
                    llm_confint[1,]+llm_confint[grepl("center", rownames(confint(llm_with_pc))),]))

llm_coefs <- cbind.data.frame(Mean = llm_coefs, llm_confint)
#
olm_coefs <- -1 * c( coef(olm_with_pc)[c("dde", "PC1")],
   "center5" = as.numeric(coef(olm_with_pc)[1]),
   coef(olm_with_pc)[1] + coef(olm_with_pc)[ grepl("center", names(coef(olm_with_pc))) ] )
olm_confint <- suppressMessages(confint(olm_with_pc))
```

95\% (approximate) confidence intervals for the models:
```{r, echo=F, include=T,cache=T}
# Will need more of a plot here
round(llm_coefs[c("dde", "PC1"),], 3)
round(llm_coefs[c("dde", "PC1"),], 3)
```

  * Hard to see improvement from a "baseline" model (Gaussian)

## Diagnostic Plots

```{r, echo=F, include=T}
par(mfrow = c(1, 2))
binnedplot(fitted(llm_with_pc),
           residuals(llm_with_pc, type = "response"),
           nclass = 40,
           xlab = "Fitted response",
           ylab = "Residual",
           main = "Logistic regression (40 bins)")
est_probs_llm <- predict(llm_with_pc, type = "response")
boxplot(est_probs_llm ~ Longnecker_complete_pc$early,
        xlab = "Preterm delivery",
        ylab = "Estimated probabilities")
#
est_probs_olm <- predict(olm_with_pc, type = "probs")
boxplot(est_probs_olm, ylab = "Estimated probabilities")
#
# DO WE WANT SOME KIND OF OUT-OF-SAMPLE PREDICTABILITY VALIDATION???
#
anova(llm_with_pc, test = "Chisq")
anova(update(llm_with_pc, . ~ . - dde + dde), test = "Chisq") 
```

## Inference on Effect Sizes
... (clear interpretation of the logistic/ordinal coefficients needed here)

## Further Discussion: PCBs

- Aggregating information of all the PCB's (PCA -> hard to interpretate)
- Selecting representative PCB's (Frequentist and Bayesian variable selection)
  - Bayesian Model Averaging
  - Horseshoe Prior
  - Hierarchical Prior
  - Lasso (Gaussian: dde, pcb_028, 074, 153; Logistic: dde, pcb_074, 153) -> 'consistent' with the previous model
  
```{r, cache=T,echo=F,eval=F}
# switch to T to include; I don't have your file!!
load("./HP_pcb.rdata")
quant5 = function(x) {round(quantile(x, c(0.025, 0.5, 0.975)),2)} ## 95% CI
pcb = dplyr::select(data.frame(beta.sim),starts_with("pcb_"))
pcb_fit<- cbind(lwr= apply(pcb, 2, quant5)[1,], fit= apply(pcb, 2, quant5)[2,], upr=apply(pcb, 2, quant5)[3,])

ggplot(as.data.frame(pcb_fit), aes(rownames(pcb_fit), 
  fit, size=10, group=1, ylim=max(upr)+0.8)) +
  theme_bw(base_size=10)+ 
  geom_point(size=2)+
  geom_line(size=1)+
  geom_errorbar(aes(ymin = lwr, ymax = upr), width = 0.2, size=1)+
  xlab("Lab")+
  ylab("Fitted value")+
  ggtitle("Gestational Age acorss (JAGS)")+
  theme(axis.text.x=element_blank())+
  geom_text(aes(label = round(lwr,2), y = lwr), vjust = 1.5, size=3) +
  geom_text(aes(label = round(upr,2), y = upr), vjust = -0.5, size=3) +
  geom_text(aes(label = rownames(pcb_fit), y = upr+0.5), vjust = -0.5, size=3, col="blue")
```


```{r lasso, cache=T,echo=F, eval=T}
library("glmnet")
data=readRDS("Longnecker.rds")
data=data%>%dplyr::select(gestational_age,dde:cholesterol,center)
data_46=data%>%mutate(y=gestational_age)
data_46[data_46$y>46,"y"]=46
data_46=data_46%>%dplyr::select(-gestational_age,-albumin)%>%mutate(race=factor(race),center=factor(center))
data_46_wscore=data_46%>%na.omit()
set.seed(123)
train.data=data_46_wscore
x <- model.matrix(y~., train.data)[,-1]
# Convert the outcome (class) to a numerical variable
y <- train.data$y

cv.lasso1 <- cv.glmnet(x, y, alpha = 1)
# Fit the final model on the training data
model1 <- glmnet(x, y, alpha = 1,
                lambda = cv.lasso1$lambda.min)

train.data=data_46_wscore
x <- model.matrix(y~., train.data)[,-1]
# Convert the outcome (class) to a numerical variable
y <- ifelse(train.data$y<37,1,0)

cv.lasso2 <- cv.glmnet(x, y, alpha = 1,family="binomial")
# Fit the final model on the training data
model2 <- glmnet(x, y, alpha = 1,family="binomial",
                lambda = cv.lasso2$lambda.min)

plot(cv.lasso1) # normal
plot(cv.lasso2) # logistic

coef(model1) # dde, pcb_028, 074, 153
coef(model2) # dde, pcb_074, 153
```

```{r vs, cache=T,echo=F}
model_n0=lm(y~.-dde-pcb_028-pcb_052-pcb_074-pcb_105-pcb_118-pcb_153-pcb_170-pcb_138-pcb_180-pcb_194-pcb_203,data=data_46_wscore)
model_ndde=lm(y~.-pcb_028-pcb_052-pcb_074-pcb_105-pcb_118-pcb_153-pcb_170-pcb_138-pcb_180-pcb_194-pcb_203,data=data_46_wscore)
model_npcbs=lm(y~.-dde-pcb_052-pcb_105-pcb_118-pcb_170-pcb_138-pcb_180-pcb_194-pcb_203,data=data_46_wscore)
data_46_wscore37=data_46_wscore
data_46_wscore37$y=ifelse(data_46_wscore37$y<37,1,0)
model_l0=glm(y~.-dde-pcb_028-pcb_052-pcb_074-pcb_105-pcb_118-pcb_153-pcb_170-pcb_138-pcb_180-pcb_194-pcb_203,data=data_46_wscore37,family = "binomial")
model_ldde=glm(y~.-pcb_028-pcb_052-pcb_074-pcb_105-pcb_118-pcb_153-pcb_170-pcb_138-pcb_180-pcb_194-pcb_203,data=data_46_wscore37,family = "binomial")
model_lpcbs=glm(y~.-dde-pcb_052-pcb_105-pcb_118-pcb_170-pcb_138-pcb_180-pcb_194-pcb_203,data=data_46_wscore37,family = "binomial")
anova(model_n0,model_ndde)
anova(model_n0,model_npcbs)
anova(model_l0,model_ldde)
anova(model_l0,model_lpcbs)
summary(model_ndde)
summary(model_npcbs)
summary(model_ldde)
summary(model_ldde)
#coef(model1) # dde, pcb_028, 074, 153
#coef(model2) # dde, pcb_074, 153
```

## Possible Improvements

  - Pooling heterogeneous effects across centers
  
  - Incorporating interactions: systematic, priors-based approach
  
  - Different methods to tackle nonlinearity
  