---
title: ""
author: "Youngsoo Baek, YunranChen and Xiaojun Zheng"
date: "1/20/2020"
output: 
  ioslides_presentation:
    widescreen: true
    smaller: true
---

<style>
slides > slide {
  overflow-x: auto !important;
  overflow-y: auto !important;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(corrplot)
library(naniar)
library(ggplot2)
library(dplyr)
library(tibble)
library(tidyr)
library(purrr)
library(lme4)
data=readRDS("Longnecker.rds")
data=data%>%select(gestational_age,dde:cholesterol,center)
```

# EDA

## EDA: Missing Data

- Over 90% missing `albumin`, around 20% missing `score_*`, others less than 0.1%
- Drop the `albumin` and keep the complete cases

```{r missing, echo=FALSE}
gg_miss_upset(data)
```

## EDA: Correlation Plot

- Weak correlation between covariates and `gestational age`
- Large correlation among covariates especially for PCBs

```{r corrplot, echo=FALSE}
data_=data%>%mutate(race=as.numeric(race))
M2=cor(as.matrix(na.omit(data_)))
corrplot.mixed(M2,number.cex=.5, upper="ellipse")
```

## EDA: Histogram for All Variables

- 0-inflation on some of PCBs
- Long left tail of `gestational age` (after truncated at 46)

```{r hist, echo=FALSE, message=FALSE, warning=FALSE}
data_46=data_#%>%mutate(y=gestational_age)
data_46[data_46$gestational_age>46,"gestational_age"]=46
data_46=data_46%>%select(-albumin)%>%mutate(race=factor(race),center=factor(center))

data_46 %>%
  keep(is.numeric) %>% 
  gather() %>% 
  ggplot(aes(value)) +
    facet_wrap(~ key, scales = "free") +
    geom_histogram()
```

## EDA: Heterogeneity across Centers

The demographic background of each center varies a lot, suggesting heterogeity across centers.
(favor fixed effect instead of random effect)

```{r}
ggplot(data_46,mapping = aes(x=as.factor(center),y=gestational_age,color=race))+geom_boxplot()+theme_bw()
ggplot(data_46,mapping = aes(x=as.factor(center),y=score_income,color=race))+geom_boxplot()+theme_bw()
```


# Challenges

## Model

## Result

## Conclusion